Validating Cell Tower Clustering for Mobility Prediction

### 2012-12-12 docs/paper589.tex
typo

The goal of the oscillation detection algorithm is to retrieve sets of towers that overlap each other -- that is, towers that could be in a stationary user's handover logs. The main idea behind the algorithm is that a user's phone is more likely to switch between oscillating towers than between regular towers that are not overlapping. In this context, \textit{switching} means a handover from one tower to another inside of a mobility path. We can define $k$ to be the minimum switching count and consider all pairs of towers that have more than $k$ switches in one mobility path to be an \textit{oscillating pair}. The example in \cite{mobilityprofiler} is illustrative: consider a sequence of handovers $\left[x, y, x, w, v, w, y\right]$ where every element represents one cell tower. We say that the pair $(x, y)$ has 3 switches: the first twofrom $x$ to $y$ and back to $x$ in the first positions of the sequence, and then back to $y$ at the end. Hence, if we adopt $k=3$ the only oscillating pair would be $(x, y)$. Note that switches do not need to be consecutive. The authors argue that this allows for detection of overlapping towers in dense region where multiple towers may be within range at the same time.

@@

As part of future work, we would like to repeat the experiments here with more data. As discussed before, only 9.44\% of the clusters in the MDC dataset had enough information to be reliably processed by the algorithm. Another avenue of research is to investigate clustering algorithms that use a different clustering quality metric. One possible approach would be the modularity metric defined by Newman and Girvan~\cite{newman}. There is also a large body of literature on clustering in theoretical computer science, especially in social networks: Andersen and Peres find sparse cuts efficently in~\cite{sparsecuts}, while Aroroa et. al try to find overlapping communities \cite{overlapping}. Finding overlapping communities has interesting applications for dealing with the oscillation effect, since towers can conceivably in more than one cluster.

>>>>

The goal of the oscillation detection algorithm is to retrieve sets of towers that overlap each other -- that is, towers that could be in a stationary user's handover logs. The main idea behind the algorithm is that a user's phone is more likely to switch between oscillating towers than between regular towers that are not overlapping. In this context, \textit{switching} means a handover from one tower to another inside of a mobility path. We can define $k$ to be the minimum switching count and consider all pairs of towers that have more than $k$ switches in one mobility path to be an \textit{oscillating pair}. The example in \cite{mobilityprofiler} is illustrative: consider a sequence of handovers $\left[x, y, x, w, v, w, y\right]$ where every element represents one cell tower. We say that the pair $(x, y)$ has 3 switches: the first two from $x$ to $y$ and back to $x$ in the first positions of the sequence, and then back to $y$ at the end. Hence, if we adopt $k=3$ the only oscillating pair would be $(x, y)$. Note that switches do not need to be consecutive. The authors argue that this allows for detection of overlapping towers in dense region where multiple towers may be within range at the same time.

@@

As part of future work, we would like to repeat the experiments here with more data. As discussed before, only 9.44\% of the clusters in the MDC dataset had enough information to be reliably processed by the algorithm. Another avenue of research is to investigate clustering algorithms that use a different clustering quality metric. One possible approach would be the modularity metric defined by Newman and Girvan~\cite{newman}. There is also a large body of literature on clustering in theoretical computer science, especially in social networks: Andersen and Peres find sparse cuts efficiently in~\cite{sparsecuts}, while Aroroa et. al try to find overlapping communities \cite{overlapping}. Finding overlapping communities has interesting applications for dealing with the oscillation effect, since towers can conceivably in more than one cluster.
